Motion planning for assessing accessibility of protein tunnels

### 2017-09-10 paper.tex

\linesnumbered
\begin{algorithm}[bh]
{\small
\setstretch{0.88}
\caption{\label{alg::main}Main loop of the RRT planner}
\KwIn{
    tunnel $T=( (c_i, r_i) )$, $i=1,\ldots,n$, with spheres centers $c_i \in \R^3$ and radii $r_i$,
    initial configuration $\qinit$
}
\KwData{
   ligand conformations $\L$,
   scale limits $\smin, \smax$ and $\sdelta$
}
\KwOut{
    configuration tree $\T$\;
}
\hrule
$v = 1$; // index of the virtual goal\\
$iteration = 0$\;
$\T$.addNode($\qinit$)\;
\While{$iteration < \Imax$ {\bf and}  $v < n$}{
    \eIf{$rand() < \gb$}{
        $\qrand$ = random sample around virtual goal $c_v\!\in\!T$\;
    }{
        $\qrand$ = random sample from $\C$\;
    }
    $\qnear$ = nearest node in $\T$ towards $\qrand$\;
    expand($\qnear,\qrand$)\;
    \For{$i= n-1,n-2,\ldots,v+1,v$}{ \nllabel{alg::main:a}
        $d$ = nearest node in the tree towards sphere $c_i$\;
        \If{$d < \dt $}{
            $v = i+1$; // new virtual goal found\;
            {\bf  break}\;
        }
    } \nllabel{alg::main:b}
    $iteration = iteration+1$\;
}
\return $\T$\;
}
\end{algorithm}

@@

%Therefore, we propose to measure the distance as the smallest 3D Euclidean distance between atoms placed at $\qrand$ and a configuration $q$ being tested.
The atom-based metric can be evaluated fast using KD-trees with complexity $\mathcal{O}(\log n)$, where $n$ is the number of atoms of the ligand.
For each conformation $\l \in \L$ of the ligand, a KD-tree is built for the ligand at the configuration $q=(0,0,0,0,0,0,l,\smax)$
(i.e., for ligand with no translation and no rotation).
To find the nearest atom to $\qrand$ for a ligand at the configuration $q=(x,y,z,r_x,r_y,r_z,l,s)$, the translation vector $T=(x,y,z)$ and
rotation matrix $R$ given by $r_x$, $r_y$, and $r_z$ are computed.
Then, $\qrand$ is translated to $\qrand' = R^{-1}(\qrand-T)$ and the KD-tree search is used to find the nearest point towards $\qrand'$, which
is also the nearest point of the ligand at $q$ to $\qrand$.

\begin{algorithm}[th]
{\small
\setstretch{0.88}
\caption{\label{alg::expand}expand}
\KwIn{
   configuration $\qnear$ to be expanded,
   random configuration $\qrand$
}
\KwData{
   ligand conformations $\L$,
   scale limits $\smin, \smax$, and $\sdelta$,
   tree $\T$
}
%\KwOut{
%    set $S$ of collision-free configurations reachable from $\qnear$\;
%}
\hrule
\ForEach{$l \in \L$}{
    \ForEach{$s \in (\smax,\smax-\sdelta, \ldots, \smin+\sdelta, \smin)$}{
        $\qnew = \emptyset$; // empty configuration \nllabel{alg::expa} \\
        \For{$i = 1,\ldots,m$}{
            $q=\qnear$\;
            $q.position$ = random 3D position around $\qnear$\;
            $q.rotation$ = random 3D rotation\;
            $q.l = l$\; 
            $q.s = s$\;
            \If{isCollisionFree($q$)}{
                \If{$\qnew = \emptyset$ {\bf or} $\da(q, \qrand) < \da(\qnew,\qrand)$}{ \nllabel{alg::expand:a}
                    $\qnew = q$\;
                }
            } 
        }
        \nllabel{alg::expb}
        \If{$\qnew \ne \emptyset$} {
            $\T$.addNode($\qnew$)\;
            $\T$.addEdge($\qnear,\qnew$)\;
            {\bf break;} // go to the next conformation
        }
    }
}
}
\end{algorithm}

@@

@@

\begin{figure}[b!]
\centering
%\includegraphics[width=0.25\textwidth]{fig/m040-conformations}
\includegraphics[width=0.14\textwidth]{fig/m040-conf1} \hskip 25pt
\includegraphics[width=0.11\textwidth]{fig/m040-conf2}
\caption{\label{fig::m040c}
 The examples of a `long' conformation (left) and a `packed' one for the ligand m040 (1,5-Dichloropentane).
}
\end{figure}

>>>>

\linesnumbered
\begin{algorithm}[bh]
{\small
\setstretch{0.88}
\caption{\label{alg::main}Main loop of the RRT planner}
\KwIn{
    tunnel $T=( (c_i, r_i) )$, $i=1,\ldots,n$, with spheres centers $c_i \in \R^3$ and radii $r_i$,
    initial configuration $\qinit$
}
\KwData{
   ligand conformations $\L$,
   scale limits $\smin, \smax$ and $\sdelta$,
   distance $\dt$ to move virtual goal
}
\KwOut{
    configuration tree $\T$\;
}
\hrule
$v = 1$; // index of the virtual goal\\
$iteration = 0$\;
$\T$.addNode($\qinit$)\;
\While{$iteration < \Imax$ {\bf and}  $v < n$}{
    \eIf{$rand() < \gb$}{
        $\qrand$ = random sample around virtual goal $c_v\!\in\!T$\;
    }{
        $\qrand$ = random sample from $\C$\;
    }
    $\qnear$ = nearest node in $\T$ towards $\qrand$\;
    expand($\T, \qnear,\qrand$)\;
    \For{$i= n-1,n-2,\ldots,v+1,v$}{ \nllabel{alg::main:a}
        $d$ = nearest node in the tree towards sphere $c_i$\;
        \If{$d < \dt $}{
            $v = i+1$; // new virtual goal found\;
            {\bf  break}\;
        }
    } \nllabel{alg::main:b}
    $iteration = iteration+1$\;
}
\return $\T$\;
}
\end{algorithm}

@@

%Therefore, we propose to measure the distance as the smallest 3D Euclidean distance between atoms placed at $\qrand$ and a configuration $q$ being tested.
The atom-based metric can be evaluated fast using KD-trees with complexity $\mathcal{O}(\log n)$, where $n$ is the number of atoms of the ligand.
For each conformation $\l \in \L$ of the ligand, a KD-tree is built for the ligand at the configuration $q=(0,0,0,0,0,0,l,\smax)$
(i.e., for ligand with no translation and no rotation).
To find the nearest atom to $\qrand$ for a ligand at the configuration $q=(x,y,z,r_x,r_y,r_z,l,s)$, the translation vector $V=(x,y,z)$ and
rotation matrix $R$ given by $r_x$, $r_y$ and $r_z$ are computed.
Then, $\qrand$ is translated to $\qrand' = R^{-1}(\qrand-V)$ and the KD-tree search is used to find the nearest point towards $\qrand'$, which
is also the nearest point of the ligand at $q$ to $\qrand$.

\begin{algorithm}[th]
{\small
\setstretch{0.88}
\caption{\label{alg::expand}expand}
\KwIn{
   tree $\T$,
   configuration $\qnear$ to be expanded,
   random configuration $\qrand$
}
\KwData{
   ligand conformations $\L$,
   scale limits $\smin, \smax$, and $\sdelta$,
}
%\KwOut{
%    set $S$ of collision-free configurations reachable from $\qnear$\;
%}
\hrule
\ForEach{$l \in \L$}{
    \ForEach{$s \in (\smax,\smax-\sdelta, \ldots, \smin+\sdelta, \smin)$}{
        $\qnew = \emptyset$; // empty configuration \nllabel{alg::expa} \\
        \For{$i = 1,\ldots,m$}{
            $q=\qnear$\;
            $q.position$ = random 3D position around $\qnear$\;
            $q.rotation$ = random 3D rotation\;
            $q.l = l$\; 
            $q.s = s$\;
            \If{isCollisionFree($q$)}{
                \If{$\qnew = \emptyset$ {\bf or} $\da(q, \qrand) < \da(\qnew,\qrand)$}{ \nllabel{alg::expand:a}
                    $\qnew = q$\;
                }
            } 
        }
        \nllabel{alg::expb}
        \If{$\qnew \ne \emptyset$} {
            $\T$.addNode($\qnew$)\;
            $\T$.addEdge($\qnear,\qnew$)\;
            {\bf break;} // go to the next conformation
        }
    }
}
}
\end{algorithm}

@@

\begin{figure}[bh]
\centering
%\includegraphics[width=0.25\textwidth]{fig/m040-conformations}
\includegraphics[width=0.14\textwidth]{fig/m040-conf1} \hskip 25pt
\includegraphics[width=0.11\textwidth]{fig/m040-conf2}
\caption{\label{fig::m040c}
 The examples of a `long' conformation (left) and a `packed' one for the ligand m040 (1,5-Dichloropentane).
}
\end{figure}

@@
